---
title: "Homework 2"
author: "Zhao, Qian"
date: "2018/2/7"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## (a)
*Proof*
\begin{align}
  l(\theta)&= \ln\prod\limits^n_{i=1}\frac{1}{\pi[1 + (x_i - \theta)^2]}\\
  &= \sum\limits^n_{i=1}\ln[\frac{1}{\pi[1 + (x_i - \theta)^2]}]\\
  &= n\ln\pi^{-1} + \sum\limits^n_{i=1}\ln[1 + (\theta - x_i)^2]^{-1}\\
  &= -n\ln\pi - \sum\limits^n_{i=1}\ln[1 + (\theta - x_i)^2].
\end{align}
$$l'(\theta) = -\sum\limits^n_{i=1}\frac{1}{1 + (\theta - x_i)^2}\times2(\theta - x_i) = -2\sum\limits^n_{i=1}\frac{\theta - x_i}{1 + (\theta - x_i)^2}$$.
We know
$$p(x) = \frac{1}{\pi(1 + x^2)},$$
and
$$p'(x) = -\frac{2x}{\pi(1 + x^2)^2},$$
then
\begin{align}
  I(\theta)&= n\int\frac{[p'(x)]^2}{p(x)}dx\\
  &= \frac{4n}{\pi}\int^{\infty}_{-\infty}\frac{x^2}{(1 + x^2)^3}dx,
\end{align}
Let $x = \tan{y}$, so $dx = \frac{1}{\cos^2{x}}$
then,
\begin{align}
  I(\theta)&= \frac{4n}{\pi}\int^{\frac{\pi}{2}}_{-\frac{\pi}{2}}\frac{\tan^2x}{(1 + \tan^2x)^3}\times\frac{1}{\cos^2x}dx\\
  &= \frac{4n}{\pi}\int^{\frac{\pi}{2}}_{-\frac{\pi}{2}}\tan^2x\cos^4xdx\\
  &= \frac{4n}{\pi}\int^{\frac{\pi}{2}}_{-\frac{\pi}{2}}\sin^2x\cos^2xdx\\
  &= \frac{4n}{\pi}\int^{\frac{\pi}{2}}_{-\frac{\pi}{2}}\sin^2x(1 - \sin^2x)dx\\
  &= \frac{4n}{\pi}(\int^{\frac{\pi}{2}}_{-\frac{\pi}{2}}\sin^2xdx - \int^{\frac{\pi}{2}}_{-\frac{\pi}{2}}\sin^4xdx)\\
  &= \frac{8n}{\pi}(\int^{\frac{\pi}{2}}_0\sin^2xdx - \int^{\frac{\pi}{2}}_0\sin^4xdx)\\
  &= \frac{8n}{\pi}\times(\frac{\pi}{4} - \frac{3\pi}{16})\\
  &= \frac{n}{2}
\end{align}

## (b)
The graph of the log-likehood function
```{r}
x <- c(1.77, -0.23, 2.76, 3.80, 3.47, 56.75, -1.34, 4.24, -2.44, 
       3.29, 3.71, -2.40, 4.53, -0.07, -1.05, -13.87, -2.53, -1.75)
n <- length(x)
theta <- seq(-3, 5, 0.1)
loglikehood <- function(theta){
  return(-n*log(pi) - sum(log(1 + (theta - x)^2)))
}
b <- lapply(theta, loglikehood) #b is l(theta)
plot(theta, b, type = 'l', main = 'Log-likehood', xlab = 'theta', ylab = 'l(theta)')
```
The MLE of $\theta$
```{r}
theta_0 <- c(-11, -1, 0, 1.5, 4, 4.7, 7, 8, 38)
Dloglikehood <- function(theta){
  return(-2*sum((theta - x)/(1 + (theta - x)^2)))
}
DDloglikehood <- function(theta){
  return(-2*sum((1 - (theta - x)^2)/(1 + (theta - x)^2)^2))
}
Newton_Raphson <- function(theta_t){
  while(abs(Dloglikehood(theta_t))>10^(-10)){
    theta_t <- theta_t - Dloglikehood(theta_t)/DDloglikehood(theta_t)
  }
  return(theta_t)
}
c <- lapply(theta_0, Newton_Raphson)
c
```
When the start point is 4, the results is the $\theta$ of MLE.
```{r}
x_mean <- mean(x)
Newton_Raphson(3.25)
```
It has the same result with start point 4, so the sample mean is a good starting point.

## (c)
```{r}
alpha <- c(1, 0.64, 0.25)
fixedpoint_1 <- function(theta_t){
  while(abs(Dloglikehood(theta_t))>10^(-10)){
    theta_t <- theta_t - Dloglikehood(theta_t)
  }
  return(theta_t)
}
fixedpoint_2 <- function(theta_t){
  while(abs(Dloglikehood(theta_t))>10^(-10)){
    theta_t <- theta_t - Dloglikehood(theta_t)*0.64
  }
  return(theta_t)
}
fixedpoint_3 <- function(theta_t){
  while(abs(Dloglikehood(theta_t))>10^(-10)){
    theta_t <- theta_t - Dloglikehood(theta_t)*0.25
  }
  return(theta_t)
}
```
The resluts went wrong.

##(d)
```{r}
Fisher_Scoring <- function(theta_t){
  while(abs(Dloglikehood(theta_t))>10^(-10)){
    theta_t <- theta_t + 2*Dloglikehood(theta_t)/n
  }
  return(theta_t)
}
d <- lapply(theta_0, Fisher_Scoring)
d
```

## (e)

The Newton method is the fastest, and the Fisher scoring is the most stable.

# Question 2
## (a)
The log-likehood function of $\theta$ is 
$$l(\theta) = \sum^n_{i=1}\ln[1-\cos(x_i - \theta)] - n\ln2\pi$$
Graph
```{r}
x <- c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96,
       2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52)
n <- length(x)
theta <- seq(-pi, pi, 0.1)
loglikehood <- function(theta){
  return(-n*log(2*pi) + sum(log(1 - cos(x - theta))))
}
b <- lapply(theta, loglikehood) #b is l(theta)
plot(theta, b, type = 'l', main = 'Log-likehood', xlab = 'theta', ylab = 'l(theta)')
```

## (b)

\begin{align}
  E[X|\theta]&= \frac{1}{2\pi}\int^{2\pi}_0x[1 - \cos(x - \theta)]dx\\
  &= \frac{1}{2\pi}[\frac{1}{2}x^2|^{2\pi}_0 - \frac{1}{2\pi}\int^{2\pi}_0x\cos(x - \theta)dx]\\
  &= \pi - \frac{1}{2\pi}\int^{2\pi}_0xd\sin(x-\theta)\\
  &= \pi - \frac{1}{2\pi}x\sin(x - \theta)|^{2\pi}_0 + \frac{1}{2\pi}\int^{2\pi}_0\sin(x-\theta)dx\\
  &= \pi + \sin\theta - \frac{1}{2\pi}\cos(x - \theta)|^{2\pi}_0\\
  &= \pi + \sin\theta = \overline x
\end{align}
So, $\hat\theta_{moment} = \arcsin(\overline x - \pi)$
```{r}
theta_m <- asin(mean(x) - pi)
theta_m
```

## (c)

```{r}

Dloglikehood <- function(theta){
  return(sum(sin(theta - x)/(1 - cos(theta - x))))
}
DDloglikehood <- function(theta){
  return(sum(1/(1 - cos(theta - x))))
}
Newton_Raphson <- function(theta){
  while(abs(Dloglikehood(theta_m))>10^-10){
    theta <- theta - (Dloglikehood(theta))/(DDloglikehood(theta))
  }
  return(theta)
}
```
The result went wrong.

## (d)

No results

## (e)

```{r}
thetaSeq <- seq(-pi, pi, by = 2*pi/199)
```


# Question 3
## (a)

```{r}
beetles <- data.frame(
  days    = c(0,  8,  28,  41,  63,  69,   97, 117,  135,  154),
  beetles = c(2, 47, 192, 256, 768, 896, 1120, 896, 1184, 1024))
t <- beetles$days
N <- beetles$beetles
K <- seq(0, 10, 0.01)
r <- seq(-5, 5, 0.01)
```


